---
title: "Practical R Markdown - Live Demo"
author: "Chris Mellinger"
date: "2/7/2018"
output: 
  pdf_document:
    toc: yes
---

# R Markdown as a Text Editor and Processor

R Markdown is an excellent text processing tool. It does inline equations, $\pi = 3.14$. It will do equations set off from the text too:

$$e = (1 + \frac{1}{n})^n$$

This example page has explanations: <http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html>. Other stuff it can do:

- Bulleted lists, like this one.
- **Bold text.**
- *Italicized text.*
- ~~strikethrough text~~
- Mult-tiered lists
    - list tier 2
        1) Numbered lists
        2) Continued
        1) That will autonumber no matter what you put on the left.
        
# Example Analysis

Let's get the packages we will need for this analysis.

```{r import packages, message=FALSE}
library(CMUtils)
library(ResultsHelper)
library(knitr)
library(papaja)
```

## Descriptives

```{r}
summary(mtcars)
```

```{r}
hist(mtcars$mpg)
```

```{r}
mtcars$trans <- ifelse(mtcars$am, 'auto', 'manual')

kable(table(mtcars$trans), col.names = c('Transmission Type', 'Count'))
```

## Engine Size and Efficiency

**Question:** What is the relationship between a car's engine size and its efficiency?

First pass with a simple regression:

```{r}
smSum <- lmSummary(simpleMod <- lm(mpg ~ disp, data=mtcars))
```

# Example Results Section

In the first step of the analysis, efficiency (in MPG) was regressed on engine displacement. Engine size sigificantly predicted efficiency, $b=`r round(coef(simpleMod)['disp'], 2)`$, $t(`r simpleMod$df.residual`) = `r round(smSum$coefficients['disp', 't value'], 2)`$.

*Same thing but better code.*

In the first step of the analysis, efficiency (in MPG) was regressed on engine displacement. Engine size sigificantly predicted efficiency, `r fpStr('disp', simpleMod)`.

*One more time!*

In the first step of the analysis, efficiency (in MPG) was regressed on engine displacement. Engine size sigificantly predicted efficiency, `r apa_print(simpleMod)$full_result$disp`.

# Code Chunk Options

```{r, eval=F}
`r fpStr('disp', simpleMod)`
```
